PyBOP: A Python package for battery model optimisation and parameterisation

### 2024-12-09 joss/paper.preprint.tex
feat: updates figures, contour plot into plotly subplots.

The Python Battery Optimisation and Parameterisation (\texttt{PyBOP})
package provides methods for parameterising and optimising battery
models, offering both deterministic and stochastic approaches with
example workflows to assist users. \texttt{PyBOP} enables parameter
identification of various battery models, including the electrochemical
and equivalent circuit models provided by the popular open-source
\texttt{PyBaMM} package (\citeproc{ref-Sulzer:2021}{Sulzer et al.,
2021}).

Similarly, \texttt{PyBOP} can be used for design optimisation under
user-defined operating conditions across a variety of model structures
and design goals. \texttt{PyBOP} facilitates model parameterisation
using a range of methods, with diagnostics for examining the performance
and convergence of the resulting parameters. Identified parameters can
be used for prediction, on-line control, and design optimisation,
supporting improved battery utilisation and development.

@@

\texttt{PyBOP} is a Python package that provides a user-friendly,
object-oriented interface for optimising battery model parameters.
\texttt{PyBOP} leverages the open-source \texttt{PyBaMM} package
(\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}) to formulate and solve
battery models. \texttt{PyBOP} is intended to serve a broad audience
including students, engineers, and researchers in academia and industry,
enabling the use of battery models for performance prediction where
previously this was not possible without specialised knowledge of both
modelling, parameter inference and software development. \texttt{PyBOP}
emphasises clear and informative diagnostics and workflows to support
users with varying expertise, and provides access to a wide range of
different optimisation and sampling algorithms. These methods are
provided through interfaces to \texttt{PINTS}
(\citeproc{ref-Clerx:2019}{Clerx et al., 2019}), \texttt{SciPy}
(\citeproc{ref-SciPy:2020}{Virtanen et al., 2020}), and \texttt{PyBOP}'s
own algorithms such as adaptive moment estimation with weight decay
(AdamW), gradient descent, and Cuckoo search.

\texttt{PyBOP} supports the battery parameter exchange (BPX) standard
(\citeproc{ref-BPX:2023}{Korotkin et al., 2023}) for sharing battery
parameter sets. These parameter sets are typically costly to obtain due
to the specialised equipment and time required for characterisation
experiments, the need for battery domain knowledge, and the
computational cost of parameter estimation. \texttt{PyBOP} reduces these
costs by providing fast parameter estimation, standardised workflows,
and parameter set interoperability (via BPX).

This package complements other lithium-ion battery modelling packages
built around \texttt{PyBaMM}, such as \texttt{liionpack} for battery
pack simulation (\citeproc{ref-Tranter2022}{Tranter et al., 2022}) and
\texttt{pybamm-eis} for numerical impedance spectroscopy, since the
identified parameters from \texttt{PyBOP} are easily exportable to these
other packages.

@@

\texttt{PyBOP} has a layered structure designed to compute forward model
predictions, process the results, and package the necessary information
for the optimisation and sampling algorithms. The forward model is
solved using the popular battery modelling package, \texttt{PyBaMM},
with construction, parameterisation, and discretisation managed by
\texttt{PyBOP}'s model interface to \texttt{PyBaMM}. This provides a
robust object construction process with a consistent interface between
models and optimisers. The statistical methods and optimisation
algorithms are constructed to interface cleanly with the forward model
predictions. Furthermore, identifiability metrics are provided with the
estimated parameters (through Hessian approximation of the cost
functions in frequentist workflows, and posterior distributions in
Bayesian workflows).

@@

\texttt{PyBOP} formulates the inference process into four key classes,
namely the model, the problem, the cost, and the optimiser/sampler, as
shown in \autoref{fig:classes}. Each of these objects represents a base
class with child classes constructing specialised functionality for
inference or optimisation workflows. The model class constructs a
\texttt{PyBaMM} forward model for a given set of model equations
provided by \texttt{PyBaMM}, initial conditions, spatial discretisation,
and numerical solver. By composing \texttt{PyBaMM} directly into
\texttt{PyBOP}, specialised models can be constructed alongside the
standard models, which can be modified, and optimally constructed for
the inference tasks. One such example is spatial rediscretisation, which
is performed when geometric parameters are optimised. In this situation,
\texttt{PyBOP} minimally rediscretises the \texttt{PyBaMM} model while
maintaining the problem, cost, and optimiser objects, providing improved
performance benefits to users. Alongside construction of the forward
model, \texttt{PyBOP}'s model class provides methods for obtaining
sensitivities from the prediction, enabling gradient-based optimisation
algorithms. This prediction, along with its corresponding sensitivities,
is provided to the problem class for processing and exception control. A
standardised data structure is then provided to the cost classes, which
then computes a distance, design, or likelihood-based metric for
optimisation. For deterministic optimisation, the optimisers minimise
the corresponding cost function or the negative log-likelihood if a
likelihood class is provided. Bayesian inference is provided by Monte
Carlo sampling classes, which accept the LogPosterior class and sample
from it using \texttt{PINTS} based Monte Carlo algorithms at the time of
submission. In the typical workflow, the classes in
\autoref{fig:classes} are constructed in sequence.

In addition to the core architecture, \texttt{PyBOP} provides several
specialised inference and optimisation processes. One such instance is
parameter inference from EIS forward model predicitions. PyBOP
discretises the EIS forward model inso sparse mass matrix form with
accompanying auto-differentiated Jacobian. These objects are then
translated into the frequency domain with a linear solution used to
compute the battery model impedance. In this situation, the forward
models are constructed within the spatial rediscretisation workflow,
allowing for geometric parameter inference from EIS forward model
predictions. Furthermore, \texttt{PyBOP} builds on the \texttt{JAX}
(\citeproc{ref-jax:2018}{Bradbury et al., 2018}) numerical solvers
provided by \texttt{PyBaMM} by providing \texttt{JAX}-based cost
functions for automatic forward model differentiation with respect to
the parameters. This functionality provides a performance improvement,
and allows users to use many other JAX-based inference packages in order
to optimise their cost function, such as \texttt{Numpyro}
(\citeproc{ref-numpyro:2019}{Phan et al., 2019}), \texttt{BlackJAX}
(\citeproc{ref-blackjax:2024}{Cabezas et al., 2024}), and \texttt{Optax}
(\citeproc{ref-optax:2020}{DeepMind et al., 2020}).

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{figures/PyBOP_components.drawio.png}
\caption{The core \texttt{PyBOP} architecture, showcasing the base class
interfaces. Each class provide direct mapping to a classical step in the
optimisation workflow. \label{fig:classes}}
\end{figure}

@@

Similarly, the current algorithms available for optimisation tasks are
presented in \autoref{tab:optimisers}. It should be noted that
\texttt{SciPy} minimize has gradient and non-gradient methods. From now
on, the point-based parameterisation and design optimisation tasks will
simply be referred to as optimisation tasks. This simplification can be
justified by examining \autoref{eqn:parameterisation} and
\autoref{eqn:design} and confirming that deterministic parameterisation
can be viewed as an optimisation task to minimise a distance-based cost
function.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4722}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2963}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2315}}@{}}
\caption{The currently supported optimisation algorithms classified by
candidate solution type, including gradient information.
\label{tab:optimisers}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
(Meta)heuristic
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
(Meta)heuristic
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Weight decayed adaptive moment estimation (AdamW) & Covariance matrix
adaptation (CMA-ES) & Particle swarm (PSO) \\
Improved resilient backpropagation (iRProp-) & Exponential natural
(xNES) & Nelder-Mead \\
Gradient descent & Separable natural (sNES) & Cuckoo search \\
SciPy minimize & SciPy differential evolution & \\
& & \\
\end{longtable}

As discussed above, \texttt{PyBOP} provides both point-based
optimisation estimates \autoref{tab:subclasses} in addition to Monte
Carlo sampling routines in order to recover distributions of likely
parameter values. The Monte Carlco methods construct a posterior
distribution on the inference parameters, which can be used for
uncertainty and practical identifiability. The individual sampler
classes are currently composed within \texttt{PyBOP} from the
\texttt{PINTS} library, with a base sampling class implemented for
interoperability and direct integration with the \texttt{PyBOP} model,
problem, and likelihood classes. The currently supported samplers are
listed in \autoref{tab:samplers}.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1810}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2155}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1379}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2069}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2586}}@{}}
\caption{Sampling methods supported by \texttt{PyBOP}, classified
according to the proposed method. \label{tab:samplers}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Adaptive
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Slicing
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Other
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Adaptive
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Slicing
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Other
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Monomial Gamma & Delayed Rejection Adaptive & Rank---Shrinking &
Differential Evolution & Metropolis Random Walk \\
No-U-Turn & Haario Bardenet & Doubling & & Emcee Hammer \\
Hamiltonian & Haario & Stepout & & Metropolis Adjusted Langevin \\
Relativistic & Rao Blackwell & & & \\
\end{longtable}

@@

In general, battery models (after spatial discretisation) can be written
in the form of a differential-algebraic system of equations:
\begin{equation}
\frac{\mathrm{d} \mathbf{x}}{\mathrm{d} t} = f(t,\mathbf{x},\mathbf{\theta}),
\label{dynamics}
\end{equation} \begin{equation}
0 = g(t, \mathbf{x}, \mathbf{\theta})
\label{algebraic}
\end{equation} \begin{equation}
\mathbf{y}(t) = h(t,\mathbf{x}),
\label{output}
\end{equation} with initial conditions \begin{equation}
\mathbf{x}(0) = \mathbf{x}_0(\mathbf{\theta}).
\label{initial_conditions}
\end{equation}

@@

The parameterisation of battery models is challenging due to the large
number of parameters that need to be identified compared to the
measurable outputs (\citeproc{ref-Andersson:2022}{Andersson et al.,
2022}; \citeproc{ref-Miguel:2021}{Miguel et al., 2021};
\citeproc{ref-Wang:2022}{Wang et al., 2022}). A complete
parameterisation often requires a stepwise identification of smaller
sets of parameters from a variety of excitations and different data sets
(\citeproc{ref-Chen:2020}{Chen et al., 2020};
\citeproc{ref-Chu:2019}{Chu et al., 2019}; \citeproc{ref-Kirk:2022}{Kirk
et al., 2023}).

A generic data fitting optimisation problem may be formulated as:
\begin{equation}
\min_{\mathbf{\theta}} ~ \mathcal{L}_{(\mathbf{y}_i)}(\mathbf{\theta}) ~~~
\textrm{subject to equations (\ref{dynamics})\textrm{-}(\ref{initial_conditions})}
\label{eqn:parameterisation}
\end{equation}

where \(\mathcal{L} : \mathbf{\theta} \mapsto [0,\infty)\) is a cost
function that quantifies the agreement between the model,
\(\mathbf{y}(t_i)\) and a sequence of measured observations
\(\hat{\mathbf{y}_i}\) measured at times \(t_i\). Within the
\texttt{PyBOP} framework, the \texttt{FittingProblem} class packages the
model output along with the measured observations, both of which are
then passed to the \texttt{FittingCost} classes for the computation of
the specific cost function. For gradient-based optimisers, the Jacobian
of the cost function with respect to the unknown parameters,
\((\frac{\partial \mathcal{L}}{\partial \theta})\) is computed for step
size and directional information.

@@

As mentioned above, \texttt{PyBOP} also provides inference and
optimisation capabilities through numerical impedance spectroscopy. This
is based on the methods presented in the \texttt{pybamm-eis} package and
allows fast impedance computation by inversion of the sparse PyBaMM mass
matrix after scaling and subtraction of the auto-differentiated Jacobian
matrix. More information about this procedure is available in
(\citeproc{ref-pybamm-eis}{Dhoot et al., n.d.}). The
\autoref{fig:impedance-landscape} below shows the numerical impedance
prediction available in \texttt{PyBOP} alongside the cost landscape
constructed for the inference task. At the time of publication,
gradient-based optimisation and sampling methods are not available when
using an impedance workflow.

@@

To avoid over complicating this example, we will continue with
identification in the time-domain; however, in general these two
forward-model prediction methods can be combined for improved system
excitation. As gradient information is available for this problem, the
choice of distance-based cost function and optimiser is not constrained.
Due to the different magnitudes of the two parameters, we apply the
logarithmic parameter transformation offered by \texttt{PyBOP}. This
transforms the optimisers search space of the optimiser to allow for a
common step size between the parameters, which is generally is not
required, but improves convergence in this problem. As a demonstration
of the parameterisation capabilities of \texttt{PyBOP},
\autoref{fig:convergence-min-max} (left) shows the rate of convergence
for each of the distance-minimising cost functions, while
\autoref{fig:convergence-min-max} (right) shows analogous results for
maximising a likelihood. The optimisation is performed with SciPy
Minimize using the gradient-based L-BFGS-B method.

@@

Next, the performance of the various optimisation algorithms is
presented by category: gradient-based in
\autoref{fig:optimiser-inference} (left), evolutionary strategies in
\autoref{fig:optimiser-inference} (middle) and (meta)heuristics in
\autoref{fig:optimiser-inference} (right) for a mean squared error cost
function. Note that the performance of the optimiser depends on the cost
environment, prior information and corresponding hyperparameters for
each specific problem.

@@

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/contour_total.png}
\caption{Cost landscape contour plot with corresponding optimisation
traces. The top row represents the gradient-based optimisers, the middle
row is the evolution-based, and the bottom row is the (meta)heuristics.
The order from left to right corresponds to the entries in
\autoref{tab:optimisers}. \label{fig:optimiser-inference}}
\end{figure}

This parameterisation task can also be approached from a Bayesian
perspective, which we will present below using \texttt{PyBOP}'s sampler
methods. The optimisation equation presented in
\autoref{eqn:parameterisation} does not represent the Bayesian parameter
identification task, and as such we introduce the Bayes theorem as,

@@

where, \(P(\theta|D)\) is the posterior and represents the probability
density function of the parameter. \(P(D|\theta)\) is the likelihood
function and assesses the parameter values alongside a noise model.
\(P(\theta)\) encapsulates the prior knowledge about the parameters, and
finally \(P(D)\) is the model evidence and acts as a normalising
constant so that the final posterior is a correctly scaled density
function. Our goal in parameter inference is to identify the parameter
values with the highest probability, which can be represented as a
point-based metric or as the posterior distribution, which provides
additional information about the uncertainty of the identified
parameters. Monte Carlo sampling methods are available to obtain this
posterior distribution. These methods sample from the posterior using a
variety of methods, including gradient-based methods such as No-U-Turn
(\citeproc{ref-NUTS:2011}{Hoffman \& Gelman, 2011}) and Hamiltonian
(\citeproc{ref-Hamiltonian:2011}{Brooks et al., 2011}), as well as
heuristic methods such as Differential Evolution
(\citeproc{ref-DiffEvolution:2006}{Braak, 2006}), and finally
conventional methods based on random sampling with rejection criteria
(\citeproc{ref-metropolis:1953}{Metropolis et al., 1953}).
\texttt{PyBOP} offers a sampling class that provides an interface to
these samplers, which are supported by the Probabilistic Inference of
Noise Time-Series (\texttt{PINTS}) package. \autoref{fig:posteriors}
below shows the sampled posterior for the synthetic workflow described
above, using an adaptive covariance-based sampler, Haario Bardenet
(\citeproc{ref-Haario:2001}{Haario et al., 2001}).

@@

>>>>

The Python Battery Optimisation and Parameterisation (\texttt{PyBOP})
package provides methods for estimating and optimising battery model
parameters, offering both deterministic and stochastic approaches with
example workflows to assist users.

\texttt{PyBOP} enables parameter identification from data for various
battery models, including both the electrochemical and equivalent
circuit models provided by the popular open-source \texttt{PyBaMM}
package (\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}). Using the
same workflow, \texttt{PyBOP} can also be used for design optimisation
under user-defined operating conditions across a variety of model
structures and design goals. \texttt{PyBOP} facilitates optimisation
using a range of methods, with diagnostics for examining optimiser
performance and convergence of the cost and corresponding parameters.
Identified parameters can be used for prediction, on-line control, and
design optimisation to accelerate research and development and improve
battery utilisation.

@@

\texttt{PyBOP} is a Python package that provides a user-friendly,
object-oriented interface for optimising battery model parameters.
\texttt{PyBOP} leverages the open-source \texttt{PyBaMM} package
(\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}) to formulate and solve
battery models. Together, these packages serve a broad audience
including students, engineers, and researchers in academia and industry,
enabling the use of advanced models where previously this was not
possible without specialised knowledge of battery modelling, parameter
inference and software development. \texttt{PyBOP} emphasises clear and
informative diagnostics and workflows to support users with varying
levels of domain expertise, and provides access to a wide range of
optimisation and sampling algorithms. These methods are provided through
interfaces to \texttt{PINTS} (\citeproc{ref-Clerx:2019}{Clerx et al.,
2019}), \texttt{SciPy} (\citeproc{ref-SciPy:2020}{Virtanen et al.,
2020}), and \texttt{PyBOP}'s own implementations of algorithms such as
adaptive moment estimation with weight decay (AdamW), gradient descent,
and Cuckoo search.

\texttt{PyBOP} supports the battery parameter exchange (BPX) standard
(\citeproc{ref-BPX:2023}{Korotkin et al., 2023}) for sharing battery
parameter sets. These parameter sets are typically costly to obtain due
to the specialised equipment and time required for characterisation
experiments, the need for battery domain knowledge, and the
computational cost of parameter estimation. \texttt{PyBOP} reduces the
requirements for the latter two categories by providing fast parameter
estimation, standardised workflows, and parameter set interoperability
(via BPX).

This package complements other lithium-ion battery modelling packages
built around \texttt{PyBaMM}, such as \texttt{liionpack} for battery
pack simulation (\citeproc{ref-Tranter2022}{Tranter et al., 2022}) and
\texttt{pybamm-eis} for fast numerical computation of the
electrochemical impedance for any battery model, since the identified
parameters from \texttt{PyBOP} are easily exportable to these other
packages.

@@

\texttt{PyBOP} has a layered structure designed to package up the
necessary functionality to compute forward predictions, process the
results, and run the optimisation and sampling algorithms. The forward
model is solved using the popular battery modelling package,
\texttt{PyBaMM}, with construction, parameterisation, and discretisation
managed by \texttt{PyBOP}'s model interface to \texttt{PyBaMM}. This
provides a robust object construction process with a consistent
interface between models and optimisers. The statistical methods and
optimisation algorithms are constructed to interface cleanly with the
forward model predictions. Furthermore, identifiability metrics are
provided with the estimated parameters (through Hessian approximation of
the cost functions in frequentist workflows, and posterior distributions
in Bayesian workflows).

@@

\texttt{PyBOP} formulates the inference process into four key classes,
namely, the model, the problem, the cost/likelihood, and the
optimiser/sampler, as shown in \autoref{fig:classes}. Each of these
objects represents a base class with child classes constructing
specialised functionality for different workflows. The model class
constructs a \texttt{PyBaMM} forward model for a given set of model
equations, initial conditions, spatial discretisation, and numerical
solver. By composing \texttt{PyBaMM} directly into \texttt{PyBOP},
specialised models can be constructed alongside the standard models,
which can be modified and purposefully constructed for different
inference tasks. One such example is spatial re-discretisation, which is
required when one or more geometric parameters are being optimised. In
this situation, \texttt{PyBOP} rebuilds the \texttt{PyBaMM} model a
minimum number of times while maintaining the problem, cost, and
optimiser objects, providing improved performance benefits to users.
Alongside construction of the forward model, \texttt{PyBOP}'s model
class provides methods for obtaining sensitivities from the prediction,
enabling gradient-based optimisation algorithms. A forward prediction,
along with its corresponding sensitivities, is provided to the problem
class for processing and exception control. A standardised data
structure is then provided to the cost classes, which then computes a
distance, design, or likelihood-based metric for optimisation. For
deterministic optimisation, the optimisers minimise the cost function or
the negative log-likelihood if a likelihood class is provided. Bayesian
inference is provided by sampler classes, which accept the LogPosterior
class and sample from it using \texttt{PINTS} based Monte Carlo
algorithms at the time of submission. In the typical workflow, the
classes in \autoref{fig:classes} are constructed in sequence.

In addition to the core architecture, \texttt{PyBOP} provides several
specialised inference and optimisation processes. One such instance is
parameter inference from electrochemical impedance spectroscopy (EIS)
simulations. PyBOP discretises the EIS forward model into a sparse mass
matrix form with accompanying auto-differentiated Jacobian. These
objects are then translated into the frequency domain with a linear
solution used to compute the battery model impedance. In this situation,
the forward models are constructed within the spatial re-discretisation
workflow, allowing for geometric parameter inference from EIS
simulations. Furthermore, \texttt{PyBOP} builds on the \texttt{JAX}
(\citeproc{ref-jax:2018}{Bradbury et al., 2018}) numerical solvers
provided by \texttt{PyBaMM} by providing \texttt{JAX}-based cost
functions for automatic forward model differentiation with respect to
the parameters. This functionality provides a performance improvement,
and allows users to use many other JAX-based inference packages in order
to optimise their cost function, such as \texttt{Numpyro}
(\citeproc{ref-numpyro:2019}{Phan et al., 2019}), \texttt{BlackJAX}
(\citeproc{ref-blackjax:2024}{Cabezas et al., 2024}), and \texttt{Optax}
(\citeproc{ref-optax:2020}{DeepMind et al., 2020}).

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{figures/PyBOP_components.drawio.png}
\caption{The core \texttt{PyBOP} architecture, showcasing the base class
interfaces. Each class provides a direct mapping to a classical step in
the optimisation workflow. \label{fig:classes}}
\end{figure}

@@

Similarly, the current algorithms available for optimisation tasks are
presented in \autoref{tab:optimisers}. It should be noted that
\texttt{SciPy} minimize comprises a range of gradient and non-gradient
methods. From now on, the point-based parameterisation and design
optimisation tasks will simply be referred to as optimisation tasks.
This simplification can be justified by examining
\autoref{eqn:parameterisation} and \autoref{eqn:design} and confirming
that deterministic parameterisation can be viewed as an optimisation
task to minimise a distance-based cost function.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4359}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2308}}@{}}
\caption{The currently supported optimisation algorithms classified by
candidate solution type, including gradient information.
\label{tab:optimisers}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
(Meta)heuristic
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
(Meta)heuristic
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Weight decayed adaptive moment estimation (AdamW) & Covariance matrix
adaptation (CMA-ES) & Particle swarm (PSO) \\
Improved resilient backpropagation (iRProp-) & Exponential natural
(xNES) & Nelder-Mead \\
Gradient descent & Separable natural (sNES) & Cuckoo search \\
SciPy minimize & SciPy differential evolution & \\
\end{longtable}

As discussed above, in addition to point-based optimisation estimates
\autoref{tab:subclasses}, \texttt{PyBOP} also provides Monte Carlo
sampling routines in order to recover distributions of likely parameter
values. The Monte Carlo methods construct a posterior distribution on
the inference parameters, which can be used to assess uncertainty and
practical identifiability. The individual sampler classes are currently
composed within \texttt{PyBOP} from the \texttt{PINTS} library, with a
base sampler class implemented for interoperability and direct
integration with the \texttt{PyBOP} model, problem, and likelihood
classes. The currently supported samplers are listed in
\autoref{tab:samplers}.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1624}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2393}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1368}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2051}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2564}}@{}}
\caption{Sampling methods supported by \texttt{PyBOP}, classified
according to the proposed method. \label{tab:samplers}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Adaptive
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Slicing
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Other
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Adaptive
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Slicing
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Other
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Monomial Gamma & Delayed Rejection Adaptive & Rank---Shrinking &
Differential Evolution & Metropolis Random Walk \\
No-U-Turn & Haario Bardenet & Doubling & & Emcee Hammer \\
Hamiltonian & Haario & Stepout & & Metropolis Adjusted Langevin \\
Relativistic & Rao Blackwell & & & \\
\end{longtable}

@@

In general, battery models (after spatial discretisation) can be written
in the form of a differential-algebraic system of equations:
\begin{equation}
\frac{\mathrm{d} \mathbf{x}}{\mathrm{d} t} = f(t,\mathbf{x},\mathbf{\theta}),
\label{dynamics}
\end{equation} \begin{equation}
0 = g(t, \mathbf{x}, \mathbf{\theta})
\label{algebraic}
\end{equation} \begin{equation}
\mathbf{y}(t) = h(t, \mathbf{x}, \mathbf{\theta}),
\label{output}
\end{equation} with initial conditions \begin{equation}
\mathbf{x}(0) = \mathbf{x}_0(\mathbf{\theta}).
\label{initial_conditions}
\end{equation}

@@

The parameterisation of battery models is challenging due to the large
number of parameters that need to be identified compared to the
measurable outputs (\citeproc{ref-Andersson:2022}{Andersson et al.,
2022}; \citeproc{ref-Miguel:2021}{Miguel et al., 2021};
\citeproc{ref-Wang:2022}{Wang et al., 2022}). A complete
parameterisation often requires a stepwise identification of smaller
sets of parameters from a variety of excitations and different data sets
(\citeproc{ref-Chen:2020}{Chen et al., 2020};
\citeproc{ref-Chu:2019}{Chu et al., 2019}; \citeproc{ref-Kirk:2022}{Kirk
et al., 2023}; \citeproc{ref-Lu:2021}{Lu et al., 2021}).

A generic data fitting optimisation problem may be formulated as:
\begin{equation}
\min_{\mathbf{\theta}} ~ \mathcal{L}_{(\hat{\mathbf{y}_i})}(\mathbf{\theta}) ~~~
\textrm{subject to equations (\ref{dynamics})\textrm{-}(\ref{initial_conditions})}
\label{eqn:parameterisation}
\end{equation}

where \(\mathcal{L} : \mathbf{\theta} \mapsto [0,\infty)\) is a cost
function that quantifies the agreement between the model output
\(\mathbf{y}(t)\) and a sequence of observations
\((\hat{\mathbf{y}_i})\) measured at times \(t_i\). Within the
\texttt{PyBOP} framework, the \texttt{FittingProblem} class packages the
model output along with the measured observations, both of which are
then passed to the cost classes for the computation of the specific cost
function. For gradient-based optimisers, the Jacobian of the cost
function with respect to the unknown parameters,
\((\frac{\partial \mathcal{L}}{\partial \theta})\) is computed for step
size and directional information.

@@

As mentioned above, \texttt{PyBOP} also provides inference and
optimisation capabilities through numerical computation of the
electrochemical impedance. This is based on the methods presented in the
\texttt{pybamm-eis} package and allows fast impedance computation by
inversion of a sparse mass matrix after scaling and subtraction of the
auto-differentiated Jacobian matrix. More information about this
procedure is available in (\citeproc{ref-pybamm-eis}{Dhoot et al.,
n.d.}). The \autoref{fig:impedance-landscape} below shows the numerical
impedance prediction available in \texttt{PyBOP} alongside the cost
landscape constructed for the inference task. At the time of
publication, gradient-based optimisation and sampling methods are not
available when using an impedance workflow.

@@

To avoid over complicating this example, we will continue with
identification in the time-domain; however, in general these two
simulation methods can be combined for improved system excitation. As
gradient information is available for this problem, the choice of
distance-based cost function and optimiser is not constrained. Due to
the difference in magnitude between the two parameters, we apply the
logarithmic parameter transformation offered by \texttt{PyBOP}. This
transforms the search space of the optimiser to allow for a common step
size between the parameters, which is generally is not required, but
improves convergence in this problem. As a demonstration of the
parameterisation capabilities of \texttt{PyBOP},
\autoref{fig:convergence-min-max} (left) shows the rate of convergence
for each of the distance-minimising cost functions, while
\autoref{fig:convergence-min-max} (right) shows analogous results for
maximising a likelihood. The optimisation is performed with SciPy
Minimize using the gradient-based L-BFGS-B method.

@@

Next, the performance of the various optimisation algorithms is
presented by category: gradient-based methods in
\autoref{fig:optimiser-inference} (left), evolutionary strategies in
\autoref{fig:optimiser-inference} (middle) and (meta)heuristics in
\autoref{fig:optimiser-inference} (right) for a mean squared error cost
function. Note that the performance of the optimiser depends on the cost
environment, prior information and corresponding hyperparameters for
each specific problem.

@@

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/contour_subplot.png}
\caption{Cost landscape contour plot with corresponding optimisation
traces. The three rows show the gradient-based optimisers (top),
evolution strategies (middle), and (meta)heuristics (bottom).
\label{fig:optimiser-inference}}
\end{figure}

This parameterisation task can also be approached from a Bayesian
perspective, which we will present below using \texttt{PyBOP}`s sampler
methods. The optimisation equation presented in
\autoref{eqn:parameterisation} does not represent the Bayesian parameter
identification task, and as such we introduce Bayes' theorem as,

@@

where, \(P(\theta|D)\) is the posterior, representing the probability
density function of the parameter. \(P(D|\theta)\) is the likelihood
function, which assesses the parameter values alongside a noise model.
\(P(\theta)\) encapsulates the prior knowledge about the parameters, and
finally \(P(D)\) is the model evidence, which acts as a normalising
constant. Our goal in parameter inference is to identify the parameter
values with the highest probability, which can be represented either by
a point-based metric or a posterior. The posterior distribution provides
additional information about the uncertainty of the identified
parameters. Monte Carlo methods are used to sample from the posterior.
The selection of Monte Carlo methods available in \texttt{PyBOP}
includes gradient-based methods such as No-U-Turn
(\citeproc{ref-NUTS:2011}{Hoffman \& Gelman, 2011}) and Hamiltonian
(\citeproc{ref-Hamiltonian:2011}{Brooks et al., 2011}), as well as
heuristic methods such as Differential Evolution
(\citeproc{ref-DiffEvolution:2006}{Braak, 2006}), and finally
conventional methods based on random sampling with rejection criteria
(\citeproc{ref-metropolis:1953}{Metropolis et al., 1953}).
\texttt{PyBOP} offers a sampler class that provides the interface to
these samplers, which are provided by the Probabilistic Inference of
Noise Time-Series (\texttt{PINTS}) package. \autoref{fig:posteriors}
below shows the sampled posterior for the synthetic workflow described
above, using an adaptive covariance-based sampler, Haario Bardenet
(\citeproc{ref-Haario:2001}{Haario et al., 2001}).

@@

\bibitem[\citeproctext]{ref-Lu:2021}
Lu, D., Scott Trimboli, M., Fan, G., Zhang, R., \& Plett, G. L. (2021).
Implementation of a physics-based model for half-cell open-circuit
potential and full-cell open-circuit voltage estimates: {P}art {II}.
Processing full-cell data. \emph{Journal of The Electrochemical
Society}, \emph{168}(7), 070533.
\url{https://doi.org/10.1149/1945-7111/ac11a5}
